---
title: "Assignment 2"
student Id: "21314174"
author: "Rajesh Dhungna"
date: "2023-10-14"
output: html_document
---

install.packages("dplyr")
install.packages("tidyr")
install.packages("lubridate")
install.packages("ggplot2")
install.packages("corrplot")

```{r}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(corrplot)
```

Task 1

```{r}
#Loading Studentmarks dataset into sdf dataframe

sdf <- read.csv("Studentmarks.csv",header=TRUE)
head(sdf)
```


```{r}
#Shape of dataframe
dim(sdf)
```



```{r}
#Structure of dataframe
str(sdf)
```


```{r}
#Checking for NA value
any(is.na(sdf))
```


```{r}
#Exploring the summary of dataframe
summary(sdf)
```

```{r}
#Creating a new column o_dob
sdf$o_dob <- sdf$dob
```



a.Caluclating the age of student into new colum age1
```{r}
#changing the datatype of dob to Date
sdf$dob <- as.Date(sdf$dob, format="%d/%m/%Y")
```



```{r}
#Creating new column age1 by take the difference between current date and dob column
sdf$age1 <- difftime(Sys.Date(), sdf$dob)
```


We got the difference in days. So we can calculate the year by dividing days by 365.
```{r}
#First changing age1 as numeric datatype and dividing age1 by 365 and rounding to nearest integer
sdf$age1 <- round(as.numeric(sdf$age1)/365)
```


```{r}
#Checking our dataframe sdf for new column age1
head(sdf)
```


b.Splitting the original dob column into date, month and year and calculating the age2 based on year column only.

```{r}
#Separating the o_dob into date, month and year column
sdf <- separate(sdf,o_dob, c("date", "month", "year"), sep="/")

```



```{r}
#changing year colum into numeric datatype
sdf$year <- as.numeric(sdf$year)
```


```{r}
#Calculating the current year
current_year <- year(Sys.Date())
```


```{r}
#Creating a new column age2 by taking the difference between current_year and year column
sdf <- mutate(sdf, year, age2=current_year-year)
```


c.Scatter plot for Stuent name versus Marks for all three years
```{r}
#Lets gather our marks by years into separate column academci_year
marks_df <- gather(sdf,key=academic_year,value=mark,c("X2020","X2021","X2022"))

```


```{r}
#Plotting the data
ggplot(marks_df, aes(x=Studentname, y=mark, fill=academic_year,group=academic_year, color=academic_year))+
geom_point(size=2.5)+
labs(title ="Student marks in three academic years",x = " Student Name", y = "Marks")
```


d.Calculating the total marks for each students 
```{r}
#Calculating the total marks 
sdf$total_marks <- (sdf$X2020+sdf$X2021+sdf$X2022)

```


filtering those who got atleast 200 marks. 
```{r}
#Creating a new dataframe with marks atleast 200
filter_marks <- filter(sdf,total_marks>=200)
```


Rearranging the dataframe in descending order based on total marks
```{r}
#Sorting value based on total_marks
filter_marks <- arrange(filter_marks, desc(total_marks))
```


Plotting a bar chart for total_marks vs Studentname in descending order
```{r}
#Bar chart for students vs totalmarks

ggplot(filter_marks, aes(x=reorder(Studentname, total_marks, decreasing= TRUE), y= total_marks)) +
  geom_bar(stat="identity", fill="orange") +
  labs(title = "Students marks from highest to lowest", x = "Student", y = "Marks")
```


Task 2
Working with movies data



```{r}
#Importing data into movie_df dataframe 
movie_df <- read.csv("Movies.csv")
```


```{r}
#Checking the first five rows
head(movie_df)
```
```{r}
#Checking the dimension of dataframe
dim(movie_df)
```

```{r}
#Checking for structure of each attribute
str(movie_df)

```


```{r}
#Checking for na values
any(is.na(movie_df))
```

```{r}
#Total number of na in dataframe
sum(is.na(movie_df))
```


```{r}
#Exploring the summary of movie_df
summary(movie_df)
```

a. Dealing with missing values
```{r}
#Checking each columns for missing values
sapply(movie_df, function(x) length(which(x=="")))
```


```{r}
#Lets create a function to extract dimension  for each missing values
missing_index <- function(x) which(x=="", arr.ind = TRUE)
i <- missing_index(movie_df)
```


```{r}
#Replacing all missing values with NA
movie_df[i] <- NA
```

```{r}
#Lets verify for any missing values
sapply(movie_df, function(x) length(which(x=="")))
```

Now that we have changed our missing value to NA, the total NA will increase in our dataframe. First lets explore the columns which have NA values again. 
```{r}
sapply(movie_df, function(x) sum(is.na(x)))
```
 

Based on our dataset, I have decided to remove the following columns which have less influence on our target variable profit. 

1.Title: Since we are interested in finding the influential factor for the profit and each movie has unique title. Hence the title may not be appropriate profit indicator in this case.


2. Color: Most of the value for Color attribute are color so this will provide less variability in our dataset.

3. Language: All the language value in our dataset are English, hence it can be less contributor to profit.


4. Year: Since this attribute is not a continuous variable and needs separate analysis using time series method, I am excluding this column in this analysis.

5. Plot_keywords: This attribute might not be relevant in deciding the profit.

6. Movie_imdb_link : This attribute might not be relevant in deciding the profit.




```{r}
#Lets create a dataframe without these columns
m_df1 <- select(movie_df,-c(Title, Color, Language, Year, Movie_imdb_link,Plot_keywords ))
```



Since every movie has unique value for each attributes that cannot be assumed. If we impute these NA values using mode for categorical values and mean for numeric values then there is a chances that a particular movie could have high profit or rated high based on these value. Hence, I have decided to omit all the row which has NA value.

```{r}
#Lets create a new data frame without NA value
m_df <- na.omit(m_df1)
```


```{r}
#Verifying for no NA in dataset
any(is.na(m_df))
```


b. Exploration part
 
bi.
```{r}
#Lets calculate the profit
m_df$profit <- m_df$Gross-m_df$Budget
```


Before plotting the data lets normalize our dataset to bring them to same scale of 0 to 1 using min-max scaler. This will make our data consistent and easy to visualize since all the value will be converge to same scale. 

```{r}
#Defining normalization function
minmax_scale <- function(x){
  (x-min(x))/(max(x)-min(x))
}
```

Since the normalization function cannot take categorical columns as argument we must have separate dataframe with numeric value only. First lets create a dataframe which consist of numeric value only.


```{r}
#Filtering column with numeric datatype from our dataset
num_m_df <- m_df[, sapply(m_df,is.numeric)]
```


Now we can call normalization function on our new dataframe num_m_df.
```{r}
#Normalized dataframe 
nor_m_df <- as.data.frame(sapply(num_m_df,minmax_scale))
```


Now we are ready to plot our data and see the relationship using regression line. 
```{r}
#scatter plot between profit and Imdb_score
ggplot(nor_m_df, aes(y=profit,x=Imdb_score)) + geom_point()+  ggtitle("Profit by IMDB Score") + geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")
```

From the above graph we can see that there is weak correlation among Imdb_score and profit. 



```{r}
#scatter plot between profit and Gross
ggplot(nor_m_df, aes(y=profit,x=Gross)) + geom_point()+  ggtitle("Profit by Gross Income") + geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")

```

From the above graph we can see that there is slight negative correlation among gross income and profit.


```{r}
#scatter plot between profit and Reviews
ggplot(nor_m_df, aes(y=profit,x=Reviews)) + geom_point()+  ggtitle("Profit by Reviews") + geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")
```

From the above graph we can see that there is weak correlation among review and profit.



```{r}
#scatter plot between profit and Movie Facebook likes
ggplot(nor_m_df, aes(y=profit,x=Movie_facebook_likes)) + geom_point()+  ggtitle("Profit by Movie Facebook likes") + geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")
```

From the above graph we can see that there is weak correlation among review and profit.







ii. Correlation between the variables.
Since our variables are mixed type(ordinal and nominal), I have used kendall method to calculate the correlation

```{r}
#Calculating the correlation between variables
cor_d <- cor(nor_m_df, method="kendall")

```



iii.

```{r}
#Lets rename our columns to make our visualization better
colnames(cor_d) <- c("reviews","duration", "dir_fb_l","act3_fb_l","act1_fb_l","gross","votes","cast_fb_l","p_face_no","budget","act2_fb_l","imdb_scr","asp_r","m_fb_l","profit")
rownames(cor_d) <- c("reviews","duration", "dir_fb_l","act3_fb_l","act1_fb_l","gross","votes","cast_fb_l","p_face_no","budget","act2_fb_l","imdb_scr","asp_r","m_fb_l","profit")
```


```{r}
#Lets make the copy of our correlation matrix
cor_d1 <- cor_d
```


```{r}
#Lets set the diagonal correlation coefficient value to NA which is correlation with self.
diag(cor_d1) <- NA

```


```{r}
#Lets change our matrix to dataframe
cor_df <- as.data.frame(cor_d1)
```



```{r}
#Listing the maximum positive correlation coefficient between profit and other variable
max(cor_df$profit, na.rm= TRUE)
```
```{r}
#Finding the index of this value
which(cor_df==max(cor_df$profit, na.rm= TRUE), arr.ind = TRUE)
```


```{r}
#Listing maximum negative correlation
min(cor_df$profit, na.rm= TRUE)
```

```{r}
#Finding the index of this value
which(cor_df==min(cor_df$profit, na.rm= TRUE), arr.ind = TRUE)
```



Lets visualize our result in graph
```{r}
#scatter plot between profit and Budget
ggplot(nor_m_df, aes(y=profit,x=Gross)) + geom_point()+  ggtitle("Profit by Gross") + geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")
```



From above two result the maximum positive correlation is between profit and gross with the value of 0.45 and maximum negative correlation is between profit and budget with the value of -0.1. 

It is obvious that higher the gross income of the movie higher the profit which is also shown by our result. Hence if we want to increase the profit we must explore the variables that have influence on the gross sales to name few Director, Actor1 etc. 


```{r}
#Listing the minimum correlation
min(abs(cor_df$profit), na.rm= TRUE)

```

```{r}
#Finding the index of this value

which(cor_df==min(abs(cor_df$profit), na.rm= TRUE), arr.ind = TRUE)
```

So the weakest correlation between profit is with duration of movie with the value of 0.0045.


```{r}
#Lets see a line graph between profit and duration
#scatter plot between profit and duration
ggplot(nor_m_df, aes(y=profit,x=Duration)) + geom_point()+  ggtitle("Profit by Duration") + geom_smooth(method=lm,  linetype="dashed", color="darkred", fill="blue")
```


From the above result we see that the duration of the movie has very less impact on the profit. This may be because most of the value lies below the range of 120 minutes which cause minimal effect on profit.

```{r}
summary(m_df$Duration)
```


Lets plot the correlation coefficient for all the variable that we have used in our analysis. 
```{r}
#Correlation Matrix plot
corrplot(cor_d,method="number")
```




```{r}
#Lets explore the value of maximum correlation coefficient in our study
max(abs(cor_df), na.rm= TRUE)
```
```{r}
#Index of maximum corelation
which(cor_df==max(abs(cor_df), na.rm= TRUE), arr.ind = TRUE)
```

This value is for the cast_facebook_likes and actor1_facebook_likes. We can interpret this result as if a lead actor of the movie have higher facebook like than the overall cast of the movie will get higher likes. 


```{r}
#Lets explore the value of minimum correlation coefficient in our study
min_correlation <- min(abs(cor_df), na.rm= TRUE)
min_correlation
```

If we search our correlation matrix this value matches -0.002067434 with number of faces on poster and movie likes on facebook. Hence we can say that there is very weak relation between Facebook like for a movie based on number of faces on poster. 


We can also visualize our correlation result using corrplot with square method for better visualization. 
```{r}
#Correlation Matrix plot
corrplot(cor_d,method="square")
```

